---
title: "analysis_script"
format: html
editor: visual
---

<!--# Hey bestie, before submitting the final version -> ctrl + f: 'NOTE' to check if all notes are evaluated and removed -->

## Version Control

```{r}
renv::restore()

library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(stringr)
library(purrr)
library(writexl)
library(openxlsx)
library(tibble)
```

## Initial simple data cleaning

```{r}
### Loading the data into R 
data <- read_excel("simulation_survey_data_direct_qualtrics_download.xlsx")

# Removing irrelevant and identifying columns:
# StartDate, EndDate, Status, RecordedDate, ResponseId, DistributionChannel
# UserLanguage
data <- data %>% select(-StartDate, -EndDate, -Status, -RecordedDate, 
                        -ResponseId, -DistributionChannel, -UserLanguage)
# Removing the first row as this contains a second header instead of data
data <- data[-1,]
```

## Participant-specific simple data cleaning

```{r}
####
## Exclude everyone that did not fill out any content-related question (Q5 and further)

# Step 1: select all columns from 15 onward (which is column where Q5 is in)
cols_from_15 <- select(data, 15:ncol(data))

# Step 2: only keep participants that have data for at least one of those selected columns (i.e, select all rows where not all selected columns are NA or -99)
data_filtered <- data %>%
  filter(rowSums(is.na(cols_from_15) | cols_from_15 == '-99') < ncol(cols_from_15))

#### (NOTE: might want to move this part to descriptives + calculate full vs partial responses)
# Calculate reponse rate
n_invitations <- 695
n_participants <- 224
response_rate <- n_participants/n_invitations # = 0.3223

####
## Add an ID variable to the dataset
data_filtered <- data_filtered %>%
  mutate(ID = paste0("id", row_number())) %>%
  relocate(ID) # put ID variable column as the first column of the dataset

####
# Change columns with values into type numeric
data_filtered <- data_filtered %>%
  mutate(across(
    where(~ is.character(.) &&
            all(str_detect(., "^\\d+$") | is.na(.) | str_detect(., "-99"))),
    as.numeric
  ))
```

## Save intermediate dataset to continue with

```{r}

data <- data_filtered

# Save data as Excel file
write.xlsx(data, file = "survey_intermediate_datafile.xlsx")
```

## Evaluating text results

In this block we will extract all textual results that need to be evaluated to decide what to do with them. The previously saved dataset can be useful to open when evaluating textual data because evaluating what to do with open answers might depend on other answers given by the same participant.

```{r}

# Define the text columns you want to extract
text_cols <- c("Q2_4_TEXT", "Q7a_12_TEXT", "Q7b_12_TEXT", 
               "Q10a_12_TEXT", "Q10b_12_TEXT", 
               "Q13a_12_TEXT", "Q13b_12_TEXT",
               "Q16a_15_TEXT", "Q16b_15_TEXT",
               "Q18a_10_TEXT", "Q18b_9_TEXT",
               "Q19_6_TEXT", "Q20_8_TEXT",
               "Q22_11_TEXT", "Q23_5_TEXT",
               "Q24_13_TEXT", "Q25_12_TEXT",
               "Q30")

# Create a named list of tibbles
text_list <- map(text_cols, function(col) {
  data_filtered %>%
    filter(.data[[col]] != "-99", !is.na(.data[[col]])) %>%
    select(ID, text = all_of(col))
}) %>%
  set_names(text_cols)

# Write to Excel with one sheet per column
#write_xlsx(text_list, "text_responses.xlsx")
```

## Anonymizing in Excel

Before going to the next step, the open answers were all anonymized by hand in Excel. Any part of an open answer that allows for an individual to be identified is replaced by '\[redacted\]'. This step does not change anything in terms of reproducibility of the other steps in this script.

## Load anonymized dataset to work with

```{r}

# Load the anonymized data here
data <- read.xlsx("survey_data_anonymized_raw.xlsx")
```

## Data Cleaning following evaluation of textual results

```{r}

####
# Removing participants we decided to exclude based on their answers

id_remove <- c("id2", "id32", "id94")

data <- data %>%
  filter(!ID %in% id_remove)


####
# Changes in individual answers based on open answer options

# Addition of new answer categories based on themes found in open answers (see codebook for details)
data <- data %>%
  add_column(Q16a_16_NEW = NA, .after = "Q16a_15_TEXT") %>% # Keyword: Not Applicable
  add_column(Q18a_11_NEW = NA, .after = "Q18a_10_TEXT") %>% # Keyword: Not Applicable
  add_column(Q16a_17_NEW = NA, .after = "Q16a_16_NEW") %>% # Keyword: Low Occurrence
  add_column(Q18a_12_NEW = NA, .after = "Q18a_11_NEW") %>% # Keyword: Uncommon
  add_column(Q18a_13_NEW = NA, .after = "Q18a_12_NEW") %>% # Keyword: Alternative Check
  add_column(Q20_9_NEW = NA, .after = "Q20_8_TEXT") %>% # Keyword: Field
  add_column(Q20_10_NEW = NA, .after = "Q20_9_NEW") %>% # Keyword: Experience
  add_column(Q20_11_NEW = NA, .after = "Q20_10_NEW") # Keyword: Not Available



####
# Impossible combinations

# Listing impossible combinations:





####
# Slider check
```

## Variable combining

```{r}

## Creating the variable 'new_method' that indicates if one of the authors has been involved in developing one or multiple methods that are part of their paper

data <- data %>%
  mutate(new_method = case_when(Q2 == "1.0" | 
                                Q2 == "2.0" | 
                                Q3 == "1.0" ~ 1, 
                                .default = 0))
```

## Save final dataset after all cleaning steps are done

```{r}

```
