---
title: "analysis_script"
format: html
editor: visual
---

## Version Control

```{r}
renv::restore()

library(tidyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(stringr)
library(purrr)
library(writexl)
library(openxlsx)
```

## Initial simple data cleaning

```{r}
### Loading the data into R 
data <- read_excel("simulation_survey_data.xlsx")

# Removing irrelevant and identifying columns:
# StartDate, EndDate, Status, RecordedDate, ResponseId, DistributionChannel
# UserLanguage
data <- data %>% select(-StartDate, -EndDate, -Status, -RecordedDate, 
                        -ResponseId, -DistributionChannel, -UserLanguage)
# Removing the first row as this contains a second header instead of data
data <- data[-1,]
```

## Participant-specific simple data cleaning

```{r}
####
## Exclude everyone that did not fill out any content-related question (Q5 and further)

# Step 1: select all columns from 15 onward (which is column where Q5 is in)
cols_from_15 <- select(data, 15:ncol(data))

# Step 2: only keep participants that have data for at least one of those selected columns (i.e, select all rows where not all selected columns are NA or -99)
data_filtered <- data %>%
  filter(rowSums(is.na(cols_from_15) | cols_from_15 == '-99') < ncol(cols_from_15))
# Response rate after this is N = 224/692 = 32.4% 

####
## Add an ID variable to the dataset
data_filtered <- data_filtered %>%
  mutate(ID = paste0("id", row_number())) %>%
  relocate(ID)
```

## Save intermediate dataset to continue with

```{r}

data <- data_filtered

# Save data as Rdata file
save(data, file = "survey_intermediate_datafile.RData")

# Save data as Excel file
write.xlsx(data, file = "survey_intermediate_datafile.xlsx")
```

## Evaluating text results

In this block we will extract all textual results that need to be evaluated to decide what to do with them. The previously saved dataset can be useful to open when evaluating textual data because evaluating what to do with open answers might depend on other answers given by the same participant.

```{r}

# Define the text columns you want to extract
text_cols <- c("Q2_4_TEXT", "Q7a_12_TEXT", "Q7b_12_TEXT", 
               "Q10a_12_TEXT", "Q10b_12_TEXT", 
               "Q13a_12_TEXT", "Q13b_12_TEXT",
               "Q16a_15_TEXT", "Q16b_15_TEXT",
               "Q18a_10_TEXT", "Q18b_9_TEXT",
               "Q19_6_TEXT", "Q20_8_TEXT",
               "Q22_11_TEXT", "Q23_5_TEXT",
               "Q24_13_TEXT", "Q25_12_TEXT",
               "Q30")

# Create a named list of tibbles
text_list <- map(text_cols, function(col) {
  data_filtered %>%
    filter(.data[[col]] != "-99", !is.na(.data[[col]])) %>%
    select(ID, text = all_of(col))
}) %>%
  set_names(text_cols)

# Write to Excel with one sheet per column
write_xlsx(text_list, "text_responses.xlsx")
```

## Load intermediate dataset to work with

```{r}

# Load the partly cleaned data here
load("survey_intermediate_datafile.RData")
```

## Data Cleaning following evaluation of textual results

```{r}

####
# Remove those participants for which an aim was entered in Q2 that does not fit the inclusion criteria

####
# Changes in individual answers based on open answer options

####
# Impossible combinations

####
# Slider check

#### 
# Other cleaning steps: make numerical values of numeric type
data <- data %>%
  mutate(across(
    where(~ is.character(.) &&
            all(str_detect(., "^\\d+$") | is.na(.))),
    as.numeric
  ))
```

## Variable combining

```{r}

## Creating the variable 'new_method' that indicates if one of the authors has been involved in developing one or multiple methods that are part of their paper

data <- data %>%
  mutate(new_method = case_when(Q2 == "1.0" | 
                                Q2 == "2.0" | 
                                Q3 == "1.0" ~ 1, 
                                .default = 0))
```

## Save final dataset after all cleaning steps are done

```{r}

```
